---
title: "Penguins"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Import and Tidy

```{r}
tuesdata <- tidytuesdayR::tt_load('2020-07-28')
penguins <- tuesdata$penguins
glimpse(penguins)
```

Let's convert chars to factors.

```{r}
penguins <- penguins %>% 
  mutate(species = factor(species), 
         island = factor(island), 
         sex = factor(sex))
glimpse(penguins)
```

# Exploration Loop 1

How does species affect bill_length_mm?

```{r}
library(ggridges)
library(patchwork)

common_theme <- theme(axis.title.y = element_blank())

length_by_species <- penguins %>% 
  ggplot(aes(bill_length_mm, species, fill = species)) +
  geom_density_ridges() +
  theme_economist() +
  labs(x = "Bill Length (mm)") +
  common_theme

depth_by_species <- penguins %>% 
  ggplot(aes(bill_depth_mm, species, fill = species)) +
  geom_density_ridges() +
  theme_economist() +
  labs(x = "Bill Depth (mm)") +
  common_theme

weight_by_species <- penguins %>% 
  ggplot(aes(flipper_length_mm, species, fill = species)) +
  geom_density_ridges() 

length_by_species / depth_by_species / weight_by_species
```
What about bill_depth_mm?

```{r}
penguins %>% 
  ggplot(aes(bill_depth_mm)) +
  geom_histogram(binwidth = 2)

penguins %>% 
  ggplot(aes(species, bill_depth_mm)) +
  geom_boxplot()
```

That's interesting. Looks like depth and length are inversely correlated. Let's confirm.

```{r}
penguins %>% 
  ggplot(aes(bill_length_mm, bill_depth_mm)) +
  geom_point() +
  facet_wrap(~ species)
```

Not really if we separate out the effect of species.

How does sex affect bill_depth and length?

```{r}
penguins %>% 
  filter(!is.na(sex)) %>% 
  ggplot(aes(sex, bill_depth_mm)) +
  geom_boxplot()

penguins %>% 
  filter(!is.na(sex)) %>% 
  ggplot(aes(sex, bill_length_mm)) +
  geom_boxplot()
```

What about body mass?

```{r}
library(ggthemes)
penguins %>% 
  filter(!is.na(sex)) %>% 
  ggplot(aes(sex, body_mass_g)) +
  geom_boxplot() +
  theme_economist()
```

Can we create a model that can predict sex based on bill_length, depth, weight, and species?

Perhaps better: can the model predict bill_length from species?

```{r}
penguins %>% 
  count(year)
```

```{r}
library(cowplot)

p <- ggplot(mtcars, aes(disp, mpg)) + 
  geom_point(size = 1.5, color = "blue") +
  theme_cowplot(12)


image <-ggplot() +
  geom_blank() +
  theme(rect = element_blank())

right <- ggdraw() + draw_image("./penguin.png")
```


```{r}
penguins %>% 
  filter(species == "Gentoo") %>% 
  ggplot(aes(flipper_length_mm)) +
  geom_histogram(binwidth = 5) +
  theme_economist()

penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(flipper_length_mm)) +
  geom_histogram(binwidth = 5) +
  theme_economist()

penguins %>% 
  filter(species == "Chinstrap") %>% 
  ggplot(aes(flipper_length_mm)) +
  geom_histogram(binwidth = 5) +
  theme_economist()
```


# Exploration Loop 2: Changes over time

```{r message = F}
library(GGally)
glimpse(penguins)
penguins %>% 
  filter(!is.na(sex)) %>% 
  ggpairs(
    aes(color = species),
    columns = c(
      "species",
      "bill_length_mm",
      "bill_depth_mm",
      "body_mass_g",
      "flipper_length_mm",
      "sex"
    ),
    columnLabels = c("species",
                     "bill length",
                     "bill depth",
                     "body mass", 
                     "flipper length",
                     "sex"))

# penguins %>% 
#   ggplot(aes(body_mass_g, flipper_length_mm)) +
#   geom_point()
```

```{r fig.asp=2}

bm_flipper_length <- penguins %>% 
  ggplot(aes(body_mass_g, flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "...longer flippers", y = "Flipper Length (mm)", x = element_blank())

bm_bill_length <- penguins %>% 
  ggplot(aes(body_mass_g, bill_length_mm, color = species)) +
  geom_point() +
  labs(title = "...longer beaks", y = "Beak Length (mm)", x = element_blank())

bm_bill_depth <- penguins %>% 
  ggplot(aes(body_mass_g, bill_depth_mm, color = species)) +
  geom_point() +
  labs(title = "...and deeper beaks", 
       subtitle = "(controlling for the impact of species)", 
       y = "Beak Depth (mm)", 
       x = "Body Mass (g)")

(bm_flipper_length / bm_bill_length / bm_bill_depth / guide_area()) +
  plot_annotation(
    title = "With these penguins, more mass often means...",
    caption = "Data from Tidy Tuesday Week 31"
    ) +
  plot_layout(guides = "collect") &
  theme_economist() &
  theme(axis.title.x = element_text(face = "bold.italic", margin = margin(t = 30)))

ggsave("my-plot.png")
```

